### 2013-06-24 doc/thesis/report.tex
THESIS DIR ADDED WITH AN INITIAL OUTLINE OF THE REPORT

\documentclass[12pt,a4paper,final]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

\author{Nacer Khalil}
\title{An internet of things testbed for the smart grid application}

\begin{document}

\begin{titlepage}
\begin{center}
\includegraphics[width=0.6\textwidth]{img/aui_logo.jpg}
%\includegraphics[width=0.45\textwidth]{img/tum_logo.jpg}
\linebreak
\linebreak
\linebreak
\linebreak
\linebreak
\linebreak
% Title
\HRule \\[0.4cm]
{ \huge \bfseries An Internet Of Things Testbed For The Smart Grid Application}\\[0.4cm]

\HRule \\[1.5cm]

% Author and supervisor
\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
Nacer \textsc{Khalil}
\end{flushleft}
\end{minipage}
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
Dr.~Riduan \textsc{Abid} \\
Dr.~Driss \textsc{Benhaddou} \\
Dr.~Michael \textsc{Gerndt} 
\end{flushright}
\end{minipage}

\vfill

% Bottom of the page
{\large \today}

\end{center}
\end{titlepage}

\chapter{Abstract}

\chapter{Acknowledgements}

%THIS IS THE TABLE OF CONTENT PART
\tableofcontents
\newpage
\chapter{Introduction}

\section{Problem statement}
\section{Purpose of study}
\section{Importance of study}
\section{scope of study}
\section{Rationale of study}
\section{Definition of terms}

\chapter{Literature review}

\chapter{Methodology}

\chapter{Findings}

\chapter{Conclusion}

\chapter{Recommendations}

\mbox{}
\nocite{*}
\bibliographystyle{IEEEtran} 
\bibliography{IEEEabrv,myrefs}
\newpage
\listoffigures
\chapter{Appendix}

\end{document}