### 2013-07-08 trunk/doc/thesis/cover_page.tex
ARABIC TITLE GRRRRRR

\begin{titlepage}
\begin{center}
\includegraphics[width=0.4\textwidth]{img/aui_logo.jpg}
% Title
\\
{ \huge \bfseries Internet of Things
\\ a Real-World Deployment Towards Smart Grids Integration}\\[1.0cm]

In partial fulfillment of the requirements for the degree of Master of Science in Software Engineering (MSSE) at the School of Science and Engineering at Al Akhawayn University in Ifrane, Morocco \\[1.0cm]

% Author and supervisor

\emph{Author:}\\
{\large \textbf{Nacer \textsc{Khalil}}}\\[1.0cm]

\emph{Supervisors:} \\
{\large \textbf{Dr.~Mohamed Riduan \textsc{Abid}}} \\
{\large \textbf{Dr.~Driss \textsc{Benhaddou}}} \\
{\large \textbf{Dr.~Michael \textsc{Gerndt}}} \\[1.0cm]

\vfill
\emph{Submitted in:}\\
{\large \textbf{Summer 2013}} \\
\end{center}
\end{titlepage}